---
title: "functions"
output: html_notebook
---

## Functions

Custom functions can be used to accomplish different operations -- or clean up code.  A standard rule of thumb is that if you have to type the same code three or more times, it's worth the effor to make that part of the code a function.  While this is certainly true for experienced coders, for the unititated coders you may spend more time writing the function. Nonetheless, functional programming is very useful.  For a more complete treatment of [functions, see the chapter](https://r4ds.had.co.nz/functions.html) in _R for Data Science_ by Wickham and Grolemund.



```{r}
make_me_a_plot <- function(df) {
  df %>% 
    ggplot(aes(height, mass)) +
    geom_point() 
}

multiply_by_two <- function(my_x) {
  mutate(my_x, doubled = height * 2)
}

foo_two <- function(my_df, my_var) {
  mutate(my_df, doubled = {{ my_var }} * 2)
}
```


```{r}
gender_subsetted_starwars %>% 
  mutate(mass_height_plot = map(data, make_me_a_plot)) %>% 
  mutate(mass_height_2x = map(data, ~ multiply_by_two(.x))) %>%
  mutate(foo = map(data, ~ foo_two(.x, height))) 
```
