---
title: "functions"
output: html_notebook
---

## Functions

Custom functions can be used to accomplish different operations -- or clean up code.  A standard rule of thumb is that if you have to type the same code three or more times, it's worth the effor to make that part of the code a function.  While this is certainly true for experienced coders, for the unititated coders you may spend more time writing the function. Nonetheless, functional programming is very useful.  For a more complete treatment of [functions, see the chapter](https://r4ds.had.co.nz/functions.html) in _R for Data Science_ by Wickham and Grolemund.

```{r}
library(tidyverse)
```


## first look

```{r}
multiplybytwo <- function(n) {
    return(2 * n)
}

plotcars <- function(df) {
  plot(df)
}

```

```{r}
multiplybytwo(9)
plotcars(cars)
```
A tidyverse way to make a function:

```{r}
make_key <- . %>% 
  mutate(key = name) %>%  
  mutate(key = str_to_lower(key)) %>%  
  mutate(key = str_extract(key, "\\w+$")) %>% 
  mutate(first_part = str_to_lower(str_extract(name, "^\\w{1,3}")))  %>% 
  mutate(key = str_glue("{key}_{first_part}")) %>% 
  select(-first_part)

starwars |>
  make_key() |>
  select(name, key)
```

The base-R way to do the same as above:

```{r}
my_key_func <- function(df) {
  df |> 
  mutate(key = name) |>  
  mutate(key = str_to_lower(key)) |>  
  mutate(key = str_extract(key, "\\w+$")) |> 
  mutate(first_part = str_to_lower(str_extract(name, "^\\w{1,3}")))  |> 
  mutate(key = str_glue("{key}_{first_part}")) |> 
  select(-first_part)
}

starwars %>% 
  my_key_func() |>
  select(name, key)
```

Or, a more complicated regex that accomplishes nearly the same goals as the custom functions above. 

```{r}
starwars |> 
  mutate(foo = str_to_lower(str_replace(name, "(^\\w{3}).*\\s(\\w+$)", "\\2_\\1"))) |>
  select(name, foo)
```

## Tidy evaluation

Read more about [Data maksing, env-variables, and indirection](https://dplyr.tidyverse.org/articles/programming.html#data-masking)

In short, you need to use curly-curly brases around the tibble's variable names when referencing those variable names within your function.

```{r}
make_me_a_plot <- function(df) {
  df %>% 
    ggplot(aes(height, mass)) +
    geom_point() 
}

make_me_a_plot(starwars)
```


```{r}
multiply_by_two <- function(my_x) {
  mutate(my_x, doubled = height * 2)
}

multiply_by_two(starwars) |>
  select(height, doubled)
```


```{r}
foo_two <- function(my_df, my_var) {
  mutate(my_df, doubled = {{ my_var }} * 2)
}

foo_two(starwars, mass) |>
  select(mass, doubled)
```

```{r}
starwars |>
  mutate(double = map_dbl(height, ~ .x * 2)) |>
  select(height, double)

map_dbl(starwars$height, ~ .x * 2)
```




```{r}
gender_subsetted_starwars %>% 
  # mutate(mass_height_plot = map(data, ~ .x %>% make_me_a_plot)) %>% 
  mutate(mass_height_2x = map(data, ~ multiply_by_two(.x))) %>%
  mutate(foo = map(data, ~ foo_two(.x, height))) 
```
