---
title: "Exercise 01"
subtitle: "Import compressed data; tidy names; unite multiple columns; separate multi-value fields; pivot long"
---

```{r}
library(tidyverse)
library(janitor)
```

1. Import "data/title.ratings.tsv.gz"
  a. What function imports tab-seperated-values files?
  b. What is the role of `..` in the code chunk below?


```{r}
my_tsv_df <- _______("../data/title.ratings.tsv.gz") 
my_tsv_df
```

Tidy the variable names with `janitor::clean_names()`

```{r}
my_tsv_df |> 
  janitor::_______()
```

Unite the `average_rating` and `num_votes` fields into `combined_variable` consisting of a forward-slash `/` value-separator.  Slice the first 100 rows and assign into a tibble named `my_united_df`

```{r}
my_united_df <- my_tsv_df |> 
  janitor::clean_names() |> 
  unite("combined_variable", c(________, _________), sep = ___) |> 
  slice(_______)
my_united_df
```

Separate the rows of the `combined_variable` in `my_united_df`.  How many rows make up this new tibble?  convert `combined_variable` into a numeric vector


```{r}
my_united_df |> 
  ___________(combined_variable) |> 
  mutate(combined_variable = _________(combined_variable))
```

In this case, the previous two operations would be simpler with `pivot_longer()`

```{r}
my_tsv_df |> 
  janitor::clean_names() |> 
  pivot_longer(cols = average_rating:num_votes, names_to = "var_type")
```

