---
title: "Exercise 02"
subtitle: "pattern matching with regex and {stringr} ; case_when()"
---

```{r}
library(tidyverse)
```

Generate a data frame by executing the next code chunk

```{r}
my_regex_df <- tribble(
  ~words,
  "eugenia005",
  "fred01",
  "lillian8372",
  "sukie64"
)

my_regex_df
```

Separate `words` into `names` and `id`.
Remember the helpful, two-page {stringr} regex cheat sheet:  https://raw.githubusercontent.com/rstudio/cheatsheets/main/strings.pdf

```{r}
my_regex_df <- my_regex_df |> 
  mutate(id = str_extract(_____, "______")) |> 
  mutate(names = str_extract(______, "______"))
my_regex_df
  
```

Convert `names` to title case

```{r}
my_regex_df |> 
  mutate(names = stringr::_________(names))
```

Want more practice with regex?  Try https://regex101.com/  Be aware the regular expressions will not be identical but similar to how they are expressed in R.  See the stringr cheatsheet to translate into R.


## case_when

Make a new variable, `height_category`, that categorizes the height of `starwars` characters into "short", "medium", and "tall".  Use the `case_when()` function.  Hint:  See the _complex combination_ example in the `dplyr::case_when()` documentation at https://dplyr.tidyverse.org/reference/case_when.html#ref-examples



```{r}
starwars |> 
  select(name, height) |> 
  mutate(height_category = case_when(
    height >  190 ~ "_____",
    height <= 167 ~ "_____",
    ______ ~ "______"
  )) 
```

